Appendix: A fragile guardrail---The minimal effect of election fraud on protest intensity in non-democracies

### 2025-11-15 Output/appendix_v4.Rmd
Edits to section on multicollinearity and on incumbent defeat by protest

# Incumbent defeat by protest

The classic model of election manipulation as an indicator of weakness would suggest that incumbents who rely more on manipulation will be more likely to fall when protest movements do emerge. The main text does not examine successful protests, which are a small minority of cases, in order to avoid measurement error in the duration models. Successful protest is, instead, examined here; there is little to suggest that incumbents who manipulate more are more vulnerable to defeat by protesters.

@@

Next, Table \@ref(tab:oust-mods) reports three logit regression models of successful protest; the dataset for these models includes all cases with a least one protest event. Model 9 is a minimal model that includes *intentional election irregularities*, *incumbent vote-share* and a control for *presidential elections*. Model 10 includes a larger number of control variables, while Model 11 adds the interaction term. As the results in the table show, greater election manipulation is not associated with a higher likelihood of incumbent defeat in any model. Figure \@ref(fig:oust-plot) shows no significant effect of election manipulation on protest success across all values of incumbent vote-share.  Taken together, this is further disconfirmation of the fraud-as-weakness model.

@@

model.oust3 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag  + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag + v2elirreg.inv , family = binomial(link = "logit"), data = protest.cases %>% filter(inc.lose_clean == "1"))
summary(model.oust3)

model.oust4 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
summary(model.oust4)

@@

# Chosen successors

@@

@@

```{r include=FALSE, echo=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

@@

```{r logit-table-small, echo=FALSE}
models2 <- list()
models2[['Model 15']] <- model.margin.binary
models2[['Model 16']] <- model.margin.binary.int

@@

```{r margin-prob-fig-small, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
plot_binary_margin_typ_n
```

```{r, include=FALSE, echo=FALSE}
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))

@@

cox2 <- coxph(Surv(d.events.num, surv_status) ~  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag +  n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
#summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]"))  #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

@@

# Using country fixed effects

>>>>

## Incumbent defeat by protest

A second possible source of selection bias in the main analysis sample is the decision to exclude cases where incumbents were defeated by protest movements. While this decision is necessary given the research design, it might still introduce bias if---for instance---many defeated incumbents manipulated heavily. The main text does not examine successful protests, which are a small minority of cases, but I include them in an analysis here. There is little to suggest that incumbents who manipulate more are more vulnerable to defeat by protesters.

@@

Next, Table \@ref(tab:oust-mods) reports three logit regression models of successful protest; the dataset for these models includes all cases with a least one protest event. Model 9 is a minimal model that includes *intentional election irregularities*, *incumbent vote-share* and a control for *presidential elections*. Model 10 includes a larger number of control variables, while Model 11 adds the interaction term. As the results in the table show, greater election manipulation is not associated with a higher likelihood of incumbent defeat in any model. Figure \@ref(fig:oust-plot) shows no significant effect of election manipulation on protest success across all values of incumbent vote-share.  Taken together, this is further evidence against the fraud-as-weakness model, and suggests the main results are not biased by excluding high-fraud cases where protesters toppled the government.

@@

model.oust3 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag  + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election +  regional_success_lag + v2elirreg.inv , family = binomial(link = "logit"), data = protest.cases %>% filter(inc.lose_clean == "1"))
summary(model.oust3)

model.oust4 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election +  regional_success_lag +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
summary(model.oust4)

@@

<!-- # Chosen successors

@@

-->

@@

```{r include=FALSE, echo=FALSE, include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

@@

```{r logit-table-small, echo=FALSE, include=FALSE}
models2 <- list()
models2[['Model 15']] <- model.margin.binary
models2[['Model 16']] <- model.margin.binary.int

@@

```{r margin-prob-fig-small, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H', include=FALSE}
plot_binary_margin_typ_n
```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))

@@

cox2 <- coxph(Surv(d.events.num, surv_status) ~  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag +  n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]"))  #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

@@

<!--
# Using country fixed effects
